---
title: "New England Warming"
author: "Adam A. Kemberling"
date: "Updated on: `r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F, 
                      comment = "", 
                      fig.align = "center")

# Packages
library(lubridate)
library(patchwork)
library(ggpmisc)
library(gt)
library(ncdf4)
library(raster)
library(heatwaveR)
library(stars)
library(rnaturalearth)
library(here)
library(sf)
library(tidyverse)
library(gmRi)

# Support Functions
source(here("R/oisst_support_funs.R"))
# source(here("R/temp_report_support.R"))

#box paths
box_paths <- research_access_paths()


# File Paths
mills_path <- box_paths$mills
res_path   <- str_replace(box_paths$res, "RES Data", "RES_Data")
okn_path   <- box_paths$okn  
oisst_path <- paste0(res_path, "OISST/oisst_mainstays/")


# Set ggplot theme for figures
theme_set(theme_bw())


# Set theme up for maps
map_theme <- list(
  theme(
    panel.border       = element_rect(color = "black", fill = NA),
    plot.background    = element_rect(color = "transparent", fill = "transparent"),
    line               = element_blank(),
    axis.title.x       = element_blank(), # turn off titles
    axis.title.y       = element_blank(),
    legend.position    = "bottom", 
    legend.title.align = 0.5))


```


**Polygons for Mapping**

```{r}

# Polygons for mapping
new_england <- ne_states("united states of america", returnclass = "sf") 
canada      <- ne_states("canada", returnclass = "sf") 
world       <- ne_countries(returnclass = "sf") 
greenland   <- ne_states(country = "greenland", returnclass = "sf") 

# Load the crop areas
gom_shape <- get_timeseries_paths("gmri_sst_focal_areas")[["apershing_gulf_of_maine"]]["shape_path"]
gom_shape <- read_sf(gom_shape)
nelme_shape <- get_timeseries_paths("nelme_regions")$NELME$shape_path
nelme_shape <- read_sf(nelme_shape)
```



**Load Satellite Data**

```{r}


# 1. Warming Rates and Rankings
# Load Warming rates/ranks
rates_path <- paste0(oisst_path, "warming_rates/annual_warming_rates")
rates_stack_all <- stack(str_c(rates_path, "1982to2020.nc"), 
                         varname = "annual_warming_rate")
ranks_stack_all <- stack(str_c(rates_path, "1982to2020.nc"), 
                         varname = "rate_percentile")


####  oisst_window_load data  ####

# What data we want to load
data_window <- data.frame(lon = c(-720, 720),
                          lat = c(-90, 90),
                          time = as.Date(c("2021-01-01", "2021-06-30")))

# why isn't 2021 loading?
t2021 <- nc_open(filename = paste0(oisst_path, "annual_observations/sst.day.mean.2021.v2.nc"))

# 2. Global SST
# Load Global Temperatures
sst_daily <- oisst_window_load(oisst_path = oisst_path, 
                               data_window = data_window, 
                               anomalies = FALSE)
sst_daily <- raster::stack(sst_daily)


# 4. Global Anomalies
# Load Global Anomalies
oisst_daily <- oisst_window_load(oisst_path = oisst_path, 
                                 data_window = data_window, 
                                 anomalies = TRUE)
oisst_daily <- raster::stack(oisst_daily)

# fix the extents that are jank for no reason
extent(sst_daily) <- extent(matrix(data = c(0, -90, 360, 90), nrow = 2))
extent(oisst_daily) <- extent(matrix(data = c(0, -90, 360, 90), nrow = 2))
```

**Make it Monthly**

```{r}

# Make it monthly
make_monthly <- function(daily_ras){
  # Months to subset with
  month_key <- str_pad(c(1:12), 2, "left", 0) %>% setNames(month.abb)

  # names to match index to
  layer_index <- names(daily_ras)
  month_index <- str_sub(layer_index, 7, 8)
  
  # Pull distinct months
  months_present <- unique(month_index)
  month_key <- month_key[which(month_key %in% months_present)]
  
  # Pull the indices that match, take means
  map(month_key, function(x){
    
    # Pull days in month
    days_in_month <- which(month_index == x)
    
    # Take mean of those days
    month_avg <- mean(daily_ras[[days_in_month]])
  }) %>% 
    setNames(names(month_key))
  
  }


# Make them monthly
sst_monthly <- make_monthly(sst_daily)
oisst_monthly <- make_monthly(oisst_daily)

# Make stars object to map
monthly_stars_sst <- map(sst_monthly, ~ st_as_stars(rotate(.x)))
monthly_stars <- map(oisst_monthly, ~ st_as_stars(rotate(.x)))
```

**Map Settings**

```{r}
# Set Labels for all the Plots
# plot_month <- "May"           # For testing single months
plot_months <- month.abb[1:6] # For Animating through each
plot_months <- setNames(plot_months, plot_months)


# Color limit
temp_limits <- c(-5, 5)
#temp_limits <- max(abs(values(oisst_monthly[[plot_month]])), na.rm = T) * c(-1,1) # Dynamic Limits

# Avg Anomaly for midpoint
color_midpoint <- 0
# color_midpoint <- mean(values(oisst_monthly[[plot_month]]), na.rm = T) # Dynamic midpoint


```

`r gmRi::use_gmri_style_rmd(css = "gmri_rmarkdown.css")`

```{r}
# Turn on the gmri font for plots
showtext::showtext_auto()
```

#  Ocean warming Patterns of New England

This report seeks to showcase how different areas in New England are warming in the larger context of the rest of the world. The two main focal areas are the Gulf of Maine and the Northeastern United States' Continental Shelf.

**What is an "Anomaly"?**

Anomalies are another way of stating the difference between an observation and the expectation based on some reference climate. In terms of sea surface temperature, the reference period used here is daily temperature from 1982-2011, and the average daily temperature is referred to as a climatology. Anomalies are then how much hotter/hotter a given day is when compared to the average temperature across that 30-year period on that given calendar day.

## Regional Temperature Timeseries {.tabset .tabset-pills}

For the following tables and graphs global satellite data has been used to generate daily sea surface temperatures for the Gulf of Maine, the Northeastern United States' continental shelf, and the global oceans. These timeseries have been made ahead of time for another project but are loaded with the following code.

**Loading Timeseries**

```{r timeseries data}

# OISST Data
gom_ts <- oisst_access_timeseries(oisst_path = box_paths$oisst_mainstays, 
                                  region_family = "gmri focus areas", 
                                  poly_name = "apershing gulf of maine")
shelf_ts <- read_csv(get_timeseries_paths("nelme_regions")$NELME$timeseries_path)
world_ts <- read_csv(str_c(oisst_path, "/global_timeseries/global_temps_oisst.csv"))


# Put them in a list 
area_list <- list(
  "Global Oceans" = world_ts,
  "Northeastern U.S. Shelf" = shelf_ts,
  "Gulf of Maine" = gom_ts)

# Make time a date
area_list <- map(area_list, ~ mutate(.x, time = as.Date(time)))


# Get Anomalies and heatwave info
area_hw <- map(area_list, pull_heatwave_events)

# Get Fahrenheit
as_farenheit <- function(x){x * (9/5) + 32}

# format columns
area_hw <- map(area_hw, function(.x){
  .x %>% mutate(
    Month = month(time, label = T, abbr = T), 
    Year = year(time),
    sst_f = as_farenheit(sst),
    seas_f = as_farenheit(seas),
    anom_f = sst_f - seas_f)
})



```

**Summarizing Data**

```{r}

# 3. Get monthly averages
area_means <- map_dfr(area_hw, function(x){
  x %>%
    group_by(Year, Month) %>% 
    summarise(avg_temp  = round(mean(sst, na.rm = T), 2),
              avg_temp_f = round(mean(sst_f, na.rm = T), 2),
              avg_anom  = round(mean(sst_anom, na.rm = T), 2),
              anom_f    = round(mean(anom_f, na.rm = T), 2),
              n_hw_days = round(sum(mhw_event, na.rm = T), 2),
              .groups = "drop") 
  
  }, .id = "Region")
```



## 2021 Anomalies & Heatwaves {.tabset .tabset-pills}

The following tables are designed to highlight the relative magnitude of the local events of the Gulf of Maine or NE Shelf with the rest of the world's oceans.

**Format Data for Tables**

```{r}
# Pivot data for horizontal Table Displays

# anomalies
anom_avgs <- area_means %>% 
  filter(Year == 2021) %>% 
  select(Region, Month, avg_anom) %>% 
  pivot_wider(names_from = Month, values_from = avg_anom)

# Same but fahreneheit
anom_avgs_f <- area_means %>% 
  filter(Year == 2021) %>% 
  select(Region, Month, anom_f) %>% 
  pivot_wider(names_from = Month, values_from = anom_f)

# heatwaves
heat_avgs <- area_means %>% 
  filter(Year == 2021) %>% 
  select(Region, Month, n_hw_days) %>% 
  pivot_wider(names_from = Month, values_from = n_hw_days)
```

### Monthly Anomalies


```{r}
# 1. Anomalies
anom_avgs %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Average Temperature Anomalies - 2021**"), 
    subtitle = paste("Degrees Celsius Above Normal")) %>%
  tab_source_note(
    source_note = md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Reference Climatolgy Period: 1982-2011.*"))


```

```{r}
# 1. Anomalies Farenheit
anom_avgs_f %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Average Temperature Anomalies - 2021**"), 
    subtitle = paste("Degrees Fahrenheit Above Normal")) %>%
  tab_source_note(
    source_note = md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Reference Climatolgy Period: 1982-2011.*"))


```


### Heatwaves Days

```{r}
# 2. HW Days
heat_avgs %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Number of Heatwave Days - 2021**")) %>%
  tab_source_note(md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Reference Climatolgy Period: 1982-2011.*"))
```

### Warming Rates



```{r warming rates}

# 4. Get warming rates for sst
area_rates <- map(area_hw, function(x){
  x <- x %>% 
    filter(between(Year, 1982, 2020)) %>% 
    group_by(Year) %>% 
    summarise(avg_temp = mean(sst, na.rm = T),
              avg_temp_f = mean(sst_f, na.rm = T))
  
  temp_lm <- lm(avg_temp ~ Year, data = x)
  temp_f_lm <- lm(avg_temp_f ~ Year, data = x)
  return(
    list(
      temp   = temp_lm,
      temp_f = temp_f_lm
    )
  )
  
})

# 5. Pull Rates
rate_table <- imap_dfr(area_rates, function(mod, area){
  tibble("Region"        = rep(area, 2),
         "Unit"          = c("Celsius", "Fahrenheit"),
         #"Intercept"     = c(coef(mod$temp)[[1]], coef(mod$anom)[[1]]),
         "Annual Change" = c(coef(mod$temp)[[2]], coef(mod$temp_f)[[2]])) %>% 
    mutate(`Annual Change` = round(`Annual Change`, 3))
})

```


```{r}

# Celsius
rate_table %>% 
  filter(Unit == "Celsius") %>% 
  select(-Unit) %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Annual Change in Sea Surface Temperature**"),
    subtitle = "Data from Years 1982-2020, Units: Celsius") %>%
  tab_source_note(md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Reference Climatolgy Period: 1982-2011.*"))


# Fahrenheit
rate_table %>% 
  filter(Unit == "Fahrenheit") %>% 
  select(-Unit) %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Annual Change in Sea Surface Temperature**"),
    subtitle = "Data from Years 1982-2020, Units: Fahrenheit") %>%
  tab_source_note(md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Reference Climatolgy Period: 1982-2011.*"))
```

## Warming Rates Comparison {.tabset}

For each area the annual increase in temperature has been calculated for the period of 1982-2020 to account for all complete years of data available with OISSTv2.

Using this data we can compare warming rates with those of other areas around the globe.

```{r}
# Clip and average the warming rates for each area
mask_ranks <- function(mask_shape){
  
  # Get stats from ranks
  m1 <- mask(rotate(ranks_stack_all), mask_shape)
  m1 <- crop(m1, mask_shape)
  rank_mean <- cellStats(m1, mean)
  rank_min <- cellStats(m1, min)
  rank_max <- cellStats(m1, max)
  
  # Get stats from rates
  m2 <- mask(rotate(rates_stack_all), mask_shape)
  m2 <- crop(m2, mask_shape)
  rate_mean <- cellStats(m2, mean)
  rate_min <- cellStats(m2, min)
  rate_max <- cellStats(m2, max)
  
  # put in table
  table_out <- tibble("Mean Rank" = rank_mean,
                "Min Rank"  = rank_min,
                "Max Rank"  = rank_max,
                "Mean Rate" = rate_mean,
                "Min Rate"  = rate_min,
                "Max Rate"  = rate_max) %>% 
    mutate_all(round, 3)
  
  # spit them out
  return(table_out)
}


# get the ranks that go with each area
gom_ranks <- mask_ranks(gom_shape)
nelme_ranks <- mask_ranks(nelme_shape)


```

### Gulf of Maine

```{r}

# Avg temp in 20__

# Rate
ne_rate   <- round(coef(area_rates$`Gulf of Maine`$temp)["Year"], 3)
ne_rate_f <- round(coef(area_rates$`Gulf of Maine`$temp_f)["Year"], 3)

# Change in Temp
change_preds_c <- predict(area_rates$`Gulf of Maine`$temp,
                          data.frame("Year" = c(1982, 2020)))
change_c <- round(change_preds_c[2] - change_preds_c[1], 2)
change_preds_f <- predict(area_rates$`Gulf of Maine`$temp_f,
                          data.frame("Year" = c(1982, 2020)))
change_f <- round(change_preds_f[2] - change_preds_f[1], 2)

# Table
gom_ranks %>% 
  mutate(Region = "Gulf of Maine") %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Rate of Warming - Global Rankings**"),
    subtitle = "Data from Years 1982-2020, Units: Celsius") %>%
  tab_source_note(md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Warming Rates Reflect Annual Change in Degrees Celsius.*"))

```

Over the period of 1982-2011, where we have data for complete years, the Gulf of Maine's average temperature has increased at a rate of `r ne_rate`${^{\circ}C/year$ or `r ne_rate_f`$^{\circ}F/year$. Increasing (`r change_c`, `r change_f`) $^{\circ}C/^{\circ}F$ over the same period.

### Northeast Shelf

```{r}

# Avg temp in 20__

# Rate
ne_rate   <- round(coef(area_rates$`Northeastern U.S. Shelf`$temp)["Year"], 3)
ne_rate_f <- round(coef(area_rates$`Northeastern U.S. Shelf`$temp_f)["Year"], 3)

# Change in Temp
change_preds_c <- predict(area_rates$`Northeastern U.S. Shelf`$temp,
                          data.frame("Year" = c(1982, 2020)))
change_c <- round(change_preds_c[2] - change_preds_c[1], 2)
change_preds_f <- predict(area_rates$`Northeastern U.S. Shelf`$temp_f,
                          data.frame("Year" = c(1982, 2020)))
change_f <- round(change_preds_f[2] - change_preds_f[1], 2)



# Table
nelme_ranks %>% 
  mutate(Region = "Northeast U.S. Continental Shelf") %>% 
  gt(rowname_col = "Region") %>% 
  tab_stubhead(label = "Region") %>% 
  tab_header(
    title = md("**Rate of Warming - Global Rankings**"),
    subtitle = "Data from Years 1982-2020, Units: Celsius") %>%
  tab_source_note(md("*Data Source: NOAA OISSTv2 Daily Sea Surface Temperature Data.*") ) %>% 
  tab_source_note(md("*Warming Rates Reflect Annual Change in Degrees Celsius.*"))
```

Over the period of 1982-2011, where we have data for complete years, the Northeast U.S. Shelf's average temperature has increased by `r ne_rate`${^{\circ}C/year$ or `r ne_rate_f`$^{\circ}F/year$. Increasing (`r change_c`, `r change_f`) $^{\circ}C/^{\circ}F$ over the same period.

## Heatwave Timelines {.tabset}

A marine heatwave is defined a when seawater temperatures exceed a seasonally-varying threshold (usually the 90th percentile) for at least 5 consecutive days. Successive heatwaves with gaps of 2 days or less are considered part of the same event.

```{r}
# Plotting Function
plot_mhw <- function(timeseries_data){
  
  
  # Set colors by name
  color_vals <- c(
    "Sea Surface Temperature" = "royalblue",
    "Heatwave Event"          = "darkred",
    "Cold Spell Event"        = "lightblue",
    "MHW Threshold"           = "gray30",
    "MCS Threshold"           = "gray30",
    "Daily Climatology"       = "gray30")
  
  
  # Set the label with degree symbol
  ylab <- expression("Sea Surface Temperature"~degree~C)
  
  
  
  # Plot the last 365 days
  p1 <- ggplot(timeseries_data, aes(x = time)) +
      geom_segment(aes(x = time, xend = time, y = seas, yend = sst), 
                   color = "royalblue", alpha = 0.25) +
      geom_segment(aes(x = time, xend = time, y = mhw_thresh, yend = hwe), 
                   color = "darkred", alpha = 0.25) +
      geom_line(aes(y = sst, color = "Sea Surface Temperature")) +
      geom_line(aes(y = hwe, color = "Heatwave Event")) +
      geom_line(aes(y = cse, color = "Cold Spell Event")) +
      geom_line(aes(y = mhw_thresh, color = "MHW Threshold"), lty = 3, size = .5) +
      geom_line(aes(y = mcs_thresh, color = "MCS Threshold"), lty = 3, size = .5) +
      geom_line(aes(y = seas, color = "Daily Climatology"), lty = 2, size = 1) +
      scale_color_manual(values = color_vals) +
      scale_x_date(date_labels = "%b", date_breaks = "1 month") +
      theme(legend.title = element_blank(),
            legend.position = "top") +
      labs(x = "", 
           y = ylab, 
           caption = paste0("Climate reference period :  1982-2011"))
    
  
  return(p1)
}

```

### World

```{r, eval = TRUE}
area_hw$`Global Oceans` %>% 
  filter(year(time) == 2021) %>% 
  plot_mhw()
```


### Northeastern US Shelf

```{r}
area_hw$`Northeastern U.S. Shelf` %>% 
  filter(year(time) == 2021) %>% 
  plot_mhw()
```

### Gulf of Maine


```{r}
area_hw$`Gulf of Maine` %>% 
  filter(year(time) == 2021) %>% 
  plot_mhw()

# when was there a cold spell
# area_hw$`Gulf of Maine` %>% 
#   filter(mcs_event == TRUE) %>% 
#   ggplot(aes(time, sst)) +
#   geom_line()
```

## Heatwave Trends {.tabset}

This section is for looking more closely at how patterns in heatwave events are progressing. For each year summaries for the number of heatwaves, their average duration, and their peak temperatures are recorded.

```{r, eval = TRUE}
#### Annual Heatwave Summary Details
hw_trend_plot <- function(hw_data, area_name){

  # prep the summary stats
  wave_summary <- hw_data %>% 
    group_by(year(time), mhw_event_no) %>% 
    summarise(total_days = sum(mhw_event, na.rm = T),
              avg_anom = mean(sst_anom, na.rm = T),
              peak_anom = max(sst_anom, na.rm = T),
              .groups = "keep") %>% 
    ungroup() %>% 
    drop_na() %>%
    group_by(`year(time)`) %>% 
    summarise(num_waves = n_distinct(mhw_event_no),
              avg_length = mean(total_days, na.rm = T),
              avg_intensity = mean(avg_anom, na.rm = T),
              peak_intensity = max(peak_anom, na.rm = T),
              .groups = "drop") %>% 
  rename(year = `year(time)`)




####  Plotting

# number of heatwaves
hw_counts <- ggplot(wave_summary, aes(y = year, x = num_waves)) +
  geom_segment(aes(yend = year, xend = 0), 
               color = gmri_cols("gmri blue")) +
  geom_point(color = gmri_cols("gmri blue")) +
  labs(x = "Heatwave Events\n(Days)", y = "")

# average duration
hw_lengths <- ggplot(wave_summary, aes(y = year, x = avg_length)) +
  geom_segment(aes(yend = year, xend = 0), 
               color = gmri_cols("orange")) +
  geom_point(color = gmri_cols("orange")) +
  labs(x = "Heatwave Duration\n(Days)", y = "")

# avg temp
hw_temps <- ggplot(wave_summary, aes(y = year, x = avg_intensity)) +
  geom_segment(aes(yend = year, xend = 0), 
               color = gmri_cols("green")) +
  geom_point(color = gmri_cols("green")) +
  labs(x = "Avg Temp. Anomaly\n(Deg C)", y = "")

# peak temp
hw_peaks <-  ggplot(wave_summary, aes(y = year, x = peak_intensity)) +
  geom_segment(aes(yend = year, xend = 0), 
               color = gmri_cols("teal")) +
  geom_point(color = gmri_cols("teal")) +
  labs(x = "Peak Temp. Anomaly\n(Deg C)", y = "")

# Line them up
side_stack <- hw_counts | hw_lengths | hw_temps | hw_peaks
side_stack <- side_stack + plot_annotation(title = area_name)

return(side_stack)
}


```


### World

```{r}
hw_trend_plot(area_hw$`Global Oceans`, "Global Oceans")
```

### Northeast Shelf

```{r}
hw_trend_plot(area_hw$`Northeastern U.S. Shelf`, "Northeastern U.S. Shelf")
```


### Gulf of Maine

```{r}
hw_trend_plot(area_hw$`Gulf of Maine`, "Gulf of Maine")
```





## Gulf of Maine 2021 Highlights {.tabset}

2021, particularly mid-May was an exceptionally hot time for the Gulf of Maine. Here is how the average temperature and anomaly strength compares across other years in the record. 

**Why flag 2012?**
In many ways 2012 was an exceptionally hot year, maintaining temperatures above marine heatwave threshold for nearly the entire year. For comparison purposes 2012 has also been flagged as it is the hottest year on recent record.

This section seeks to place the monthly temperatures in the context of previous years for the purpose of answering:

 - Is this the hottest month on record?   
 - If not, when was it hotter?   
 - If so, when was the runner up?   
 - Top 5?  
 
 
 **Summary Prep**
 
```{r}
# Pick Area
area_comp <- area_hw$`Gulf of Maine`

# Get monthly metric, across years
month_comps <- area_comp %>% 
  filter(Month %in% month.abb[1:6]) %>% 
  group_by(Year, Month) %>% 
  summarise(
    avg_temp = mean(sst),
    avg_temp_f = mean(sst_f),
    avg_anom = mean(sst_anom),
    peak_anom = max(sst_anom),
    smallest_anom = min(sst_anom),
    n_hw_days = sum(mhw_event),
    deg_over = sum(sst_anom),
    .groups = "drop") %>% 
  mutate(Year = as.character(Year))

# Get the overall averages to place mean in context:
month_overalls <- area_comp %>% 
  filter(Month %in% month.abb[1:6]) %>% 
  group_by(Month) %>% 
  summarise(
    avg_temp = mean(sst),
    avg_temp_f = mean(sst_f),
    avg_anom = mean(sst_anom),
    peak_anom = max(sst_anom),
    smallest_anom = min(sst_anom),
    n_hw_days = sum(mhw_event),
    deg_over = sum(sst_anom),
    .groups = "drop") %>% 
  mutate(Year = "All Years")

# re-join
month_comps <- bind_rows(month_comps, month_overalls)%>% 
  mutate(
    color_flag = case_when(
    Year == "2012" ~ "2012",
    Year == "2021" ~ "2021", 
    Year == "All Years" ~ "Overall Avg.",
    TRUE ~ "Remaining Years"))
```
 


### Average Temperature

The average monthly temperature is fairly straightforward, and is useful for grounding these metrics in physical units rather than "anomalies".

```{r}
# color scheme
color_key <- c("2012" = as.character(gmri_cols("orange")),
               "2021" = as.character(gmri_cols("gmri blue")),
               "Overall Avg." = "gray30",
               "Remaining Years" = "gray70")

# Plot temperature(s)
cels <- ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = avg_temp, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
              size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
              size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "Overall Avg."), size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Average Temperature"~~degree~C),
       y = "Month",
       color = "Data Source") +
  theme(legend.position = "none")

# fahrenheit
far <- ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = avg_temp_f, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
              size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
              size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "Overall Avg."), size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Average Temperature"~~degree~F),
       y = "Month",
       color = "Data Source")


(cels | far) + plot_annotation(
  title = "Gulf of Maine",
  subtitle = "Sea Surface Temperature")

```

### Average Anomaly

The average daily anomaly sets the benchmark for how far from the climate average a particular month is.

```{r}

# Plot Anomaly Strength
ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = avg_anom, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
             size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
             size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "Overall Avg."), size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Average Temperature Anomaly"~~degree~C),
       y = "Month",
       color = "Data Source",
       title = "Gulf of Maine",
       subtitle = "Average Monthly Anomaly")
```

### Anomaly Strength

Peak daily anomalies are indicative of acute thermal stress and may put species over their thermal stress limits if they are unable move/adapt to avoid/cope with them.

```{r}
# Plot Anomaly Strength
ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = peak_anom, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
             size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
             size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Highest Daily Anomaly "~~degree~C),
       y = "Month",
       color = "Data Source",
       title = "Gulf of Maine",
       subtitle = "Peak Anomaly Intensity")
```

### Smallest Anomaly

The smallest or lowest anomaly temperature is a useful indication of temperature relief, or a break from thermal stress. A high minimum anomaly value is indicative that even "troughs" in temperature may still be stressful.

```{r}
# Plot Anomaly Strength
ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = smallest_anom, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
             size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
             size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Lowest Daily Anomaly "~~degree~C),
       y = "Month",
       color = "Data Source",
       title = "Gulf of Maine",
       subtitle = "Minimum Temperature Stress")
```

### Total Temperature Distance

The additive total of daily anomalies is a proxy that tracks the excess amount of temperature in the system. 

```{r}
# Plot Anomaly Strength
ggplot(filter(month_comps, color_flag == "Remaining Years"), 
       aes(y = fct_rev(Month), x = deg_over, color = color_flag)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_point(data = filter(month_comps, color_flag == "2012"), 
             size = 3, shape = 17) +
  geom_point(data = filter(month_comps, color_flag == "2021"), 
             size = 3, shape = 17) +
  scale_color_manual(values = color_key) +
  labs(x = expression("Cumulative Degree-Distance from Climate "~~degree~C),
       y = "Month",
       color = "Data Source",
       title = "Gulf of Maine",
       subtitle = "Degrees * Days Away from Average")
```

## Regional Maps 

For the Gulf of Maine and the Northeast Shelf a specific polygon has been used to both outline the area of interest and define what data to use for any of the area-specific calculations. These have been overlayed on the maps in blue to signify what areas are which.



### World {.tabset .tabset-pills}


```{r}
# Plot global map
plot_world <- function(plot_month){
  
  # text formatting for labels
  plot_lab_sst <-  str_c("Mean Temperature - ", plot_month, " - 2021")
  plot_lab <-  str_c("Monthly Temperature Anomalies - ", plot_month, " - 2021")
  guide_lab <- expression("Temperature Anomaly"~~degree~C)
  sst_lab <- expression("Sea surface Temperature"~~degree~C)
  
  
  # SST - C
  p1 <- ggplot() +
    geom_stars(data = monthly_stars_sst[[plot_month]]) +
    geom_sf(data = world, fill = "gray90", size = .25) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent") +
    guides("fill" = guide_colorbar(title = sst_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    coord_sf(expand = F) +
    map_theme +
    labs(subtitle = plot_lab_sst)
  
  
  # Anomalies
  p2 <- ggplot() +
    geom_stars(data = monthly_stars[[plot_month]]) +
    geom_sf(data = world, fill = "gray90", size = .25) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent", 
                         limit = temp_limits, oob = scales::squish) +
    guides("fill" = guide_colorbar(title = guide_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    coord_sf(expand = F) +
    map_theme +
    labs(subtitle = plot_lab)
  
  return((p1| p2))
}

# build list
world_maps <- map(plot_months, plot_world)

```


#### Jan

```{r}
world_maps$Jan
```


#### Feb

```{r}
world_maps$Feb
```

#### Mar

```{r}
world_maps$Mar
```

#### Apr

```{r}
world_maps$Apr
```

#### May

```{r}
world_maps$May
```

#### Jun

```{r}
world_maps$Jun
```


### Northeastern US Shelf {.tabset .tabset-pills}

```{r}
# Map of Northeastern US Shelf

plot_shelf <- function(plot_month){
  
  # text formatting for labels
   # text formatting for labels
  plot_lab_sst <-  str_c("Mean Temperature - ", plot_month, " - 2021")
  plot_lab <-  str_c("Monthly Temperature Anomalies - ", plot_month, " - 2021")
  guide_lab <- expression("Temperature Anomaly"~~degree~C)
  sst_lab <- expression("Sea surface Temperature"~~degree~C)
 
  p1 <- ggplot() +
    geom_stars(data = monthly_stars_sst[[plot_month]]) + 
    geom_sf(data = new_england, fill = "gray90", size = .25) +
    geom_sf(data = canada, fill = "gray90", size = .25) +
    geom_sf(data = greenland, fill = "gray90", size = .25) +
    geom_sf(data = nelme_shape, fill = "transparent", size = .5, 
            color = gmri_cols("gmri blue")) +
    scale_x_continuous(breaks = seq(-180, 180, 5)) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent") +
    guides("fill" = guide_colorbar(title = sst_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    map_theme +
    coord_sf(xlim = c(-76.5, -64), ylim = c(35, 45.25))  +
    labs(subtitle = plot_lab_sst)
  
  

  p2 <- ggplot() +
    geom_stars(data = monthly_stars[[plot_month]]) + 
    geom_sf(data = new_england, fill = "gray90", size = .25) +
    geom_sf(data = canada, fill = "gray90", size = .25) +
    geom_sf(data = greenland, fill = "gray90", size = .25) +
    geom_sf(data = nelme_shape, fill = "transparent", size = .5, 
            color = gmri_cols("gmri blue")) +
    scale_x_continuous(breaks = seq(-180, 180, 5)) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent", 
                         limit = temp_limits, oob = scales::squish) +
    guides("fill" = guide_colorbar(title = guide_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    map_theme +
    coord_sf(xlim = c(-76.5, -64), ylim = c(35, 45.25))  +
    labs(subtitle = plot_lab)


  return((p1 | p2))
}


# build list
shelf_maps <- map(plot_months, plot_shelf)
```

#### Jan

```{r}
shelf_maps$Jan
```


#### Feb

```{r}
shelf_maps$Feb
```

#### Mar

```{r}
shelf_maps$Mar
```

#### Apr

```{r}
shelf_maps$Apr
```

#### May

```{r}
shelf_maps$May
```

#### Jun

```{r}
shelf_maps$Jun
```

### Gulf of Maine {.tabset .tabset-pills}

```{r}
# Map of Gulf of Maine
plot_gom <- function(plot_month){

  # text formatting for labels
  plot_lab_sst <-  str_c("Mean Temperature - ", plot_month, " - 2021")
  plot_lab <-  str_c("Monthly Temperature Anomalies - ", plot_month, " - 2021")
  guide_lab <- expression("Temperature Anomaly"~~degree~C)
  sst_lab <- expression("Sea surface Temperature"~~degree~C)
 
  p1 <- ggplot() +
    geom_stars(data = monthly_stars_sst[[plot_month]]) + 
    geom_sf(data = new_england, fill = "gray90", size = .25) +
    geom_sf(data = canada, fill = "gray90", size = .25) +
    geom_sf(data = greenland, fill = "gray90", size = .25) +
    geom_sf(data = gom_shape, fill = "transparent", size = .5, 
            color = gmri_cols("gmri blue")) +
    scale_x_continuous(breaks = seq(-180, 180, 5)) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent") +
    guides("fill" = guide_colorbar(title = sst_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    map_theme +
    coord_sf(xlim = c(-72.5, -64), ylim = c(39, 45.25))  +
    labs(subtitle = plot_lab_sst)
  
  p2 <- ggplot() +
    geom_stars(data = monthly_stars[[plot_month]]) + 
    geom_sf(data = new_england, fill = "gray90", size = .25) +
    geom_sf(data = canada, fill = "gray90", size = .25) +
    geom_sf(data = greenland, fill = "gray90", size = .25) +
    geom_sf(data = gom_shape, fill = "transparent", size = .5, 
            color = gmri_cols("gmri blue")) +
    scale_x_continuous(breaks = seq(-180, 180, 5)) +
    scale_fill_distiller(palette = "RdBu", na.value = "transparent", 
                         limit = temp_limits, oob = scales::squish) +
    guides("fill" = guide_colorbar(title = guide_lab, 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +  
    map_theme +
    coord_sf(xlim = c(-72.5, -64), ylim = c(39, 45.25))  +
    labs(subtitle = plot_lab)

  return((p1| p2))

}
# build list
gom_maps <- map(plot_months, plot_gom)
```

#### Jan

```{r}
gom_maps$Jan
```


#### Feb

```{r}
gom_maps$Feb
```

#### Mar

```{r}
gom_maps$Mar
```

#### Apr

```{r}
gom_maps$Apr
```

#### May

```{r}
gom_maps$May
```

#### Jun

```{r}
gom_maps$Jun
```


## Disclaimers
 
 All data views are temporary unless noted otherwise. Data for June 2021 is currently preliminary data, and may change as QA/QC procedures take place.

`r insert_gmri_footer()`

