---
title: "New Normal - Shifting Climate Baselines"
author: "Adam A. Kemberling"
date: "Updated on: `r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F, 
                      comment = "", 
                      fig.align = "center")

# Packages
library(lubridate)
library(ncdf4)
library(raster)
library(rnaturalearth)
library(sf)
library(stars)
library(gmRi)
library(here)
library(janitor)
library(knitr)
library(patchwork)
library(tidyverse)
library(kableExtra)
library(heatwaveR)
library(thematic)
library(gt)



# Support Functions
source(here("R/oisst_support_funs.R"))
source(here("R/temp_report_support.R"))

# Box paths
box_paths <- research_access_paths()


# Set ggplot theme for figures
theme_set(theme_bw())


# Set theme up for maps
map_theme <- list(
  theme(
    panel.border       = element_rect(color = "black", fill = "transparent"),
    plot.background    = element_rect(color = "transparent", fill = "transparent"),
    line               = element_blank(),
    axis.title.x       = element_blank(), # turn off titles
    axis.title.y       = element_blank(),
    legend.position    = "bottom", 
    legend.title.align = 0.5)) 

# lindsays palette:
customRdBu <- colorRampPalette(c("#67001f","#d6604d","#e0e0e0","#4393c3","#053061"))


# Polygons for mapping
new_england <- ne_states("united states of america") %>% st_as_sf(crs = 4326) 
canada      <- ne_states("canada") %>% st_as_sf(crs = 4326)
world       <- ne_countries() %>% st_as_sf(crs = 4326)
greenland   <- ne_states(country = "greenland") %>% st_as_sfc(crs = 4326)

# Base Map
base_map <- ggplot() +
  geom_sf(data = new_england, fill = "gray80", size = 0.2) +
  geom_sf(data = canada, fill = "gray80", size = 0.2) +
  geom_sf(data = greenland, fill = "gray80", size = 0.2) 
```


`r gmRi::use_gmri_style_rmd(css = "gmri_rmarkdown.css")`

```{r}
# Turn on thematic for GMRI styling of plots:
# Set Style options
# Turn on thematic:
# Set Style options
# thematic_on(
#   bg = NA,
#   fg = NA,
#   accent = NA,
#   font = "auto"
# )

# Turn on the gmri font for plots - doesn't really connect to gmri font
showtext::showtext_auto()


```


#  Shifting Norms - Moving Climate Norms to the New Climatology Period

```{r loading climatologies}
# load average global climatologies as annual averages
yr_avgs <- str_c(box_paths$oisst, "daily_climatologies/yearly_means/")
clim_82_avg    <- stack(str_c(yr_avgs, "avg_clim_1982to2011.nc "))
clim_91_avg    <- stack(str_c(yr_avgs, "avg_clim_1991to2020.nc"))
clim_shift_avg <- stack(str_c(yr_avgs, "avg_clim_shift_82to91baselines.nc"))


# climatology_lists
grid_list <- list(
  "1982-2011"      = clim_82_avg,
  "1991-2020"      = clim_91_avg,
  "Climate Shift"  = clim_shift_avg)

# rotate them
grid_list <- map(grid_list, raster::rotate)

```


```{r plotting shifts}
# convert annual_avgs to stars objects
clim_82_st    <- st_as_stars(grid_list$`1982-2011`)
clim_91_st    <- st_as_stars(grid_list$`1991-2020`) 
clim_shift_st <- st_as_stars(grid_list$`Climate Shift`) 


# Set the crs to world molleweide
world_moll      <- st_crs("+proj=moll")

# Warp to world projection
clim_82_moll    <- warp_grid_projections(clim_82_st, projection_crs = "world moll")
clim_91_moll    <- warp_grid_projections(clim_91_st, projection_crs = "world moll")
clim_shift_moll <- warp_grid_projections(clim_shift_st, projection_crs = "world moll")


# Get temp limits that fit both
overall_max <- max(c( max(clim_82_st[[1]], na.rm = T) , max(clim_91_st[[1]], na.rm = T) ))
overall_min <- max(c( min(clim_82_st[[1]], na.rm = T) , min(clim_91_st[[1]], na.rm = T) ))
temp_limits <- c(overall_min, overall_max)


# Map the two periods
plot_period <- function(in_stars_obj, period_label, crs_projection){
  period_plot <- ggplot() +
    geom_stars(data = in_stars_obj) +
    geom_sf(data = st_transform(world, crs_projection), fill = "gray80", size = 0.2) +
    scale_fill_distiller(palette = "RdYlBu", na.value = "transparent", limit = temp_limits) +
    guides("fill" = guide_colorbar(title = expression("Average Annual Sea Surface Temperature"~~degree~C),
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(4, "in"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +
    coord_sf(crs = crs_projection) +
    map_theme +
    labs(title = period_label)
  
  return(period_plot)
}
  
```


```{r, fig.width=10, fig.height=8}
# Plotting the two periods
plot_82 <- plot_period(clim_82_moll, period_label = "1982-2011 Climatology", crs_projection = world_moll)
plot_91 <- plot_period(clim_91_moll, period_label = "1991-2020 Climatology", crs_projection = world_moll)

# Plotting them stacked
stacked_clims <- (plot_82 + theme(legend.position = "none")) / plot_91
stacked_clims
```



```{r}

####  Shift in Climatology Plot

# Set diverging limits for scale
shift_limits <- max(abs(clim_shift_st[[1]]), na.rm = T) * c(-1,1)

# Plot the shift in the two
plot_shift <- ggplot() +
    geom_stars(data = clim_shift_moll) +
    geom_sf(data = st_transform(world, world_moll), fill = "gray80", size = 0.2) +
    scale_fill_distiller(palette = "RdYlBu", na.value = "transparent", limit = shift_limits) +
    guides("fill" = guide_colorbar(title = expression("Sea Surface Temperature Difference"~~degree~C),
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(4, "in"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +
    map_theme +
    coord_sf(crs = world_moll) +
    labs(title =  "Global Shifts in Annual SST from Climatology Transition")

```




```{r, fig.width=10}
# Plotting the change
plot_shift
```



## Region Extents {.tabset .tabset-pills}

The following figures look specifically at two regions. 1.) The Gulf of Maine, and 2.) the Northeastern United States' Continental Shelf. Both of these areas have been mapped below for clarity on what data is included for any regional metric.


```{r, fig.width=10}
# load shapes/timeseries paths
gmri_areas    <- get_timeseries_paths(region_group = "gmri sst focal areas")
nelme_regions <- get_timeseries_paths(region_group = "nelme regions")

# get region shapes
gom          <- read_sf(gmri_areas$apershing_gulf_of_maine$shape_path)
ne_shelf     <- read_sf(nelme_regions$NELME$shape_path)

# lat/lon limits
map_lims <- list(x = c(-78, -63), y = c(35, 47))


# Gulf of Maine
gom_map <- base_map +
  geom_sf(data = gom, 
          alpha = 0.4,
          color = gmri_cols("gmri blue"),
          fill = gmri_cols("gmri blue")) +
  scale_y_continuous(breaks = seq(30, 50, by = 1)) +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  coord_sf(xlim = map_lims$x, ylim = map_lims$y) +
  map_theme +
  labs(title = "Gulf of Maine")

#NE Shelf
shelf_map <- base_map +
  geom_sf(data = ne_shelf, alpha = 0.4,
          color = gmri_cols("teal"),
          fill = gmri_cols("teal")) +
  scale_y_continuous(breaks = seq(30, 50, by = 1)) +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  coord_sf(xlim = map_lims$x, ylim = map_lims$y) +
  map_theme +
  labs(title = "Northeast Shelf")


# Mapping Them together
gom_map | shelf_map
```



## Change in Annual Regional Temperatures {.tabset}

```{r, fig.height=4}
# masking function
mask_sf <- function(in_ras, sf_mask){
  r1 <- crop(in_ras, sf_mask)
  r2 <- mask(r1, sf_mask)
  return(r2)}

# Mask them all
masked_grids <- map(grid_list, function(in_ras){
  
  # Get means
  gom_masked    <- mask_sf(in_ras, gom) 
  shelf_masked  <- mask_sf(in_ras, ne_shelf)
  
  # return both as list
  return(list(
    "Gulf of Maine" = gom_masked,
    "Northeast Shelf" = shelf_masked
  ))
  
})
```

### Gulf of Maine Climate Shift

```{r}
# Pull Stars Object
gom_shift_st <- st_as_stars(masked_grids$`Climate Shift`$`Gulf of Maine`)

# Plot the shift in Climate
ggplot() +
  geom_stars(data = gom_shift_st) +
  geom_sf(data = new_england, fill = "gray80", size = 0.2) +
  geom_sf(data = canada, fill = "gray80", size = 0.2) +
  geom_sf(data = greenland, fill = "gray80", size = 0.2) +
  scale_y_continuous(breaks = seq(30, 50, by = 1)) +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  coord_sf(xlim = c(-71.25, -65), ylim = c(41, 44.5)) +
  scale_fill_distiller(palette = "RdYlBu", na.value = "transparent", limit = shift_limits) +
    guides("fill" = guide_colorbar(title = expression("Sea Surface Temperature Difference"~~degree~C),
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(4, "in"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +
    map_theme +
    labs(title = "Gulf of Maine Difference in Annual Climate")
```


###  Northeast Shelf Climate Shift

```{r}
# Pull Stars Object
ne_shift_st <- st_as_stars(masked_grids$`Climate Shift`$`Northeast Shelf`)

# Plot the shift in Climate
ggplot() +
  geom_stars(data = ne_shift_st) +
  geom_sf(data = new_england, fill = "gray80", size = 0.2) +
  geom_sf(data = canada, fill = "gray80", size = 0.2) +
  geom_sf(data = greenland, fill = "gray80", size = 0.2) +
  scale_y_continuous(breaks = seq(30, 50, by = 1)) +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  coord_sf(xlim = map_lims$x, ylim = map_lims$y) +
  scale_fill_distiller(palette = "RdYlBu", na.value = "transparent", limit = shift_limits) +
    guides("fill" = guide_colorbar(title = expression("Sea Surface Temperature Difference"~~degree~C),
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(4, "in"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black")) +
    map_theme +
    labs(title = "Northeast Shelf Difference in Annual Climate")
```

###  Annual Changes

```{r, fig.height = 2}

# Get Average Temp and Change in Climate
shift_table <- imap_dfr(masked_grids, function(in_ras, climatology){
  # Get means
  gom_mean <- in_ras[["Gulf of Maine"]] %>% cellStats(mean, na.rm = T)
  ne_mean  <- in_ras[["Northeast Shelf"]] %>% cellStats(mean, na.rm = T)
  
  # build table
  grid_shift <- data.frame(
    Climatology = rep(climatology, 2),
    region = c("Gulf of Maine", "Northeast Shelf"),
    avg_temp = c(gom_mean, ne_mean))
  
  return(grid_shift) }) %>% 
    mutate(Climatology = factor(Climatology, levels = c("Climate Shift", "1991-2020", "1982-2011")))


# Plot the shifts
shift_table %>% 
  filter(Climatology == "Climate Shift") %>% 
  ggplot(aes(x = avg_temp, y = region)) +
  geom_segment(aes(xend = 0, yend = region, color = region)) +
  geom_point(aes(color = region), size = 2) +
  geom_label(aes(x = avg_temp, y = region, label = round(avg_temp, 2))) +
  scale_color_gmri() +
  geom_vline(xintercept = 0, color = "gray30", linetype = 3) +
  labs(y = "", 
       x = expression("Average Annual Temperature"~~degree~C),
       color = "") +
  theme(legend.position = "bottom")
```



## Changes to Seasonality {.tabset .tabset-pills}

**Loading Climatology and Heatwave Info**

```{r}
# Load regional timeseries
gom_ts      <- read_csv(gmri_areas$apershing_gulf_of_maine$timeseries_path, col_types = cols()) %>% 
  mutate(time = as.Date(time))
ne_shelf_ts <- read_csv(nelme_regions$NELME$timeseries_path, col_types = cols()) %>%
  mutate(time = as.Date(time))


# Side by side - Heatwave Detection
# Run climatology for both periods, merge and compare
hw_comparison <- function(region_timeseries){
  
  # Run heatwave detection using new climate period
  heatwaves_82 <- pull_heatwave_events(region_timeseries, 
                                       threshold = 90, 
                                       clim_ref_period = c("1982-01-01", "2011-12-31")) %>% 
    select(time, sst, seas_82 = seas, anom_82 = sst_anom, hw_thresh_82 = mhw_thresh, cs_thresh_82 = mcs_thresh,
           hwe_82 = mhw_event, cse_82 = mcs_event, hwe_no_82 = mhw_event_no, cse_no_82 = mcs_event_no)
           
  
  
  # Run heatwave detection using new climate period
  heatwaves_91 <- pull_heatwave_events(region_timeseries, 
                                       threshold = 90, 
                                       clim_ref_period = c("1991-01-01", "2020-12-31")) %>% 
    select(time, sst, seas_91 = seas, anom_91 = sst_anom, hw_thresh_91 = mhw_thresh, cs_thresh_91 = mcs_thresh,
           hwe_91 = mhw_event, cse_91 = mcs_event, hwe_no_91 = mhw_event_no, cse_no_91 = mcs_event_no)
  
  
  # Subtract old from the new - set up flat date for plots
  base_date <- as.Date("2000-01-01")
  heatwaves_out <- left_join(heatwaves_82, heatwaves_91, by = c("time", "sst")) %>% 
    mutate(anom_shift = anom_91 - anom_82,
           clim_shift  = seas_91 - seas_82,
           upper_shift = hw_thresh_91 - hw_thresh_82,
           lower_shift = cs_thresh_91 - cs_thresh_82,
           hwe_change = case_when( 
             hwe_82 == T & hwe_91 == F ~ "No Longer HW",
             hwe_82 == F & hwe_91 == T ~ "New HW",
             hwe_82 == T & hwe_91 == T ~ "Still HW",
             hwe_82 == F & hwe_91 == F ~ "Still No HW",
             T ~ "missing case"),
           shift_direction = ifelse(clim_shift >= 0, "Temperature Increase", "Temperature Decrease"),
           year = year(time),
           yday = yday(time),
           flat_date = as.Date(yday-1, origin = base_date))
  
  
  return(heatwaves_out)
}


# Run comparisons
gom_comp <- hw_comparison(region_timeseries = gom_ts) 
ne_comp  <- hw_comparison(region_timeseries = ne_shelf_ts)


```


**Plotting Seasonality**


```{r}
# Plotting the seasonal variations
seasonality_check <- function(comp_data){
  
  # Pull climatology
  seasonal_pattern <- comp_data %>% 
    filter(between(time, as.Date("2019-01-01"), as.Date("2019-12-31"))) %>% 
    distinct(flat_date, .keep_all = T)
  
  # Plot 82 clim
  p_82 <-ggplot(seasonal_pattern, aes(x = time)) + 
    geom_line(aes(y = seas_82), 
              color = gmri_cols("gmri blue"),
              size = 1) +
    geom_line(aes(y = hw_thresh_82), linetype = 2, color = "gray60") +
    geom_line(aes(y = cs_thresh_82), linetype = 2, color = "gray60") +
    labs(x = "", y = expression("SST"~~degree~C),
         subtitle = "1982-2011 Climatology") +
    scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0))
  
  # Plot 91 Clim
  p_91 <- ggplot(seasonal_pattern, aes(x = time)) + 
    geom_line(aes(y = seas_91), 
              color = gmri_cols("teal"),
              size = 1) +
    geom_line(aes(y = hw_thresh_91), linetype = 2, color = "gray60") +
    geom_line(aes(y = cs_thresh_91), linetype = 2, color = "gray60") +
    labs(x = "", y = expression("SST"~~degree~C),
         subtitle = "1991-2020 Climatology") +
    scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0))
  
  
  # # Plot both lines together
  # p_both <- ggplot(seasonal_pattern, aes(x = time)) + 
  #   geom_ribbon(aes(ymin = seas_82, ymax = seas_91), fill = "gray80") + 
  #   geom_line(aes(y = seas_82), 
  #             color = "1982-2011 Climatology",
  #             size = .75) +
  #   geom_line(aes(y = seas_91), 
  #             color = "1991-2020 Climatology",
  #             size = .75) +
  #   scale_colour_manual(
  #     values = c("1982-2011 Climatology" = as.character(gmri_cols("gmri blue")),
  #                "1991-2020 Climatology" = as.character(gmri_cols("teal")))) +
  #   labs(x = "", y = expression("SST"~~degree~C),
  #        subtitle = "Expected Temperature from Climatologies") +
  #   scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0))
  
  
  # Plot the difference
  p_diff <- ggplot(seasonal_pattern, aes(x = time, y = clim_shift)) + 
    geom_segment(aes(yend = 0, xend = time, color = shift_direction)) +
    scale_color_gmri(reverse = T) +
    labs(x = "", 
         y = expression("Temperature Shift"~~degree~C),
         color = "Shift in Climate Norm",
         subtitle = "Difference in Climatologies") +
    scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0)) +
    guides(color = guide_legend(title.position = "top", title.hjust = 0.5)) +
    theme(legend.position = "bottom")
  
  return(
    list(
      "clim82" = p_82,
      "clim91" = p_91,
      #"clim_both" = p_both,
      "clim_shift" = p_diff)
  )
}
```


### Gulf of Maine

```{r, fig.height=8}
# Return the plots
gom_seasonality <- seasonality_check(gom_comp)

# Assemble stack
(gom_seasonality[[1]] + labs(title = "Gulf of Maine")) / gom_seasonality[[2]] / gom_seasonality[[3]]

```


### Northeast Shelf


```{r, fig.height=8}
# Return the plots
ne_seasonality <- seasonality_check(ne_comp)

# Assemble stack
(ne_seasonality[[1]] + labs(title = "Northeast Shelf")) / ne_seasonality[[2]] / ne_seasonality[[3]]

```

## Heatwave Events {.tabset .tabset-pills}

**Heatmap of Marine Heatwave Events**

```{r}
# Plotting how anomalies are now suppressed
heatwave_checks <- function(comp_data){
  
  # Set palette limits to center it on 0 with scale_fill_distiller
  limit <- max(abs(comp_data$anom_91)) * c(-1,1)
  
  # Base date for rectangle fill for NA
  base_date <- as.Date("2000-01-01")
  
  # Assemble heatmap plot
  heatwave_heatmap <- ggplot(comp_data, aes(x = flat_date, y = year)) +
    
    # background box fill for missing dates
    geom_rect(xmin = base_date, 
              xmax = base_date + 365, 
              ymin = min(comp_data$year) - .5,
              ymax = max(comp_data$year) + .5, 
              fill = "gray30", 
              color = "transparent") +
    
    # Tile for sst colors
    geom_tile(aes(fill = anom_91)) +
    
    # Points for heatwave events
    geom_point(data = filter(comp_data, (hwe_82 == TRUE | hwe_91 == TRUE)),
               aes(x = flat_date, 
                   y = year,
                   color = hwe_change), 
               size = .25)  +
    
    # Axis Labels
    scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0)) +
    scale_y_continuous(limits = c(1980.5, 2021.5), expand = c(0,0)) +
   
    
    # Colors
    scale_fill_distiller(palette = "RdYlBu", limit = limit) + # for actual anomalies
    scale_color_manual(values = c("white", "black")) + 
    
    #5 inches is default rmarkdown height for barheight
    guides("fill" = guide_colorbar(title = "Anomaly from 91-2020 Climatology", 
                                   title.position = "top", 
                                   title.hjust = 0.5,
                                   barwidth = unit(3, "inches"), 
                                   frame.colour = "black", 
                                   ticks.colour = "black"),
           color = guide_legend(title = "Change in Heatwave Record", 
                                title.position = "top", 
                                title.hjust = 0.5, 
                                override.aes = list(size = 3))) +  
     labs(x = "", 
          y = "") +
    theme_classic() +
    theme(legend.position = "bottom",
          panel.border = element_rect(color  = "black", fill = "transparent", size = 1))


  # Assemble pieces
  return(heatwave_heatmap)
  
}
```



### Gulf of Maine

```{r, fig.height=8}
heatwave_checks(gom_comp) + 
  labs(title = "Shifting Heatwave Baseline - Gulf of Maine")
```


### Northeast Shelf

```{r, fig.height=8}
heatwave_checks(ne_comp) + 
  labs(title = "Shifting Heatwave Baseline - Northeast Shelf")
```


## Heatwave Characteristics {.tabset .tabset-pills}

**Characteristics for Each Climatology**

```{r}


####  Summarize Yearly Heatwave Totals and Characteristics
hw_characteristics <- function(comparison_data, clim_group){
  
  # Switch between 81 and 92
  event_number <- switch(clim_group,
                         "82" = sym("hwe_no_82"),
                         "91" = sym("hwe_no_91"))
  
  event_flag <- switch(clim_group,
                       "82" = sym("hwe_82"),
                       "91" = sym("hwe_91"))
  
  sst_anom <- switch(clim_group,
                     "82" = sym("anom_82"),
                     "91" = sym("anom_91"))
  
  clim_period <- switch(clim_group,
                        "82" = "1982-2011",
                        "91" = "1991-2020")
    
   # 1. Characterize each heatwave event: 
  # number of heatwaves
  # average heatwave duration
  # remove NA as a distinct heatwave number
  wave_events <- comparison_data %>% 
      mutate(yr = year(time)) %>% 
      group_by(yr, !!event_number) %>% 
      summarise(total_days = sum(!!event_flag, na.rm = T),
                avg_anom   = mean(!!sst_anom, na.rm = T),
                peak_anom  = max(!!sst_anom, na.rm = T),
                .groups = "keep") %>% 
      ungroup() %>% 
      drop_na()
  
  # 2. Get Annual Means:
  wave_summary <- wave_events %>%
      group_by(yr) %>% 
      summarise(num_waves      = n_distinct(!!event_number),
                avg_length     = mean(total_days, na.rm = T),
                avg_intensity  = mean(avg_anom, na.rm = T),
                peak_intensity = max(peak_anom, na.rm = T),
                .groups = "keep") %>% 
    ungroup() %>% 
    mutate(clim = clim_period)
  
  return(wave_summary)
  
}
```


**Process Summaries**

```{r}



#### Annual Heatwave Summary Details
gom_hw_82 <- hw_characteristics(comparison_data = gom_comp, clim_group = "82")
ne_hw_82  <- hw_characteristics(comparison_data = ne_comp, clim_group = "82")
gom_hw_91 <- hw_characteristics(comparison_data = gom_comp, clim_group = "91")
ne_hw_91  <- hw_characteristics(comparison_data = ne_comp, clim_group = "91")
```


**Shift in Heatwave Characteristics**


```{r}

# Join the two periods together for climatology shift plotting, doesn't overwrite
join_hw_summs <- function(summ_82, summ_91){ # Put the data together
  summ_91 <- select(summ_91, 
                    yr,
                    w = num_waves,
                    x = avg_length,
                    y = avg_intensity,
                    z = peak_intensity) 
  
  # Need to join to preserve all years
  both_details <- summ_82 %>% 
    select(-clim) %>% 
    full_join(summ_91, by = "yr") %>% 
    mutate(w = ifelse(is.na(w), 0, w),
           x = ifelse(is.na(x), 0, x),
           y = ifelse(is.na(y), 0, y),
           z = ifelse(is.na(z), 0, z))
  return(both_details)
}


# Run the summary for both areas
gom_shift <- join_hw_summs(gom_hw_82, gom_hw_91)
ne_shift <- join_hw_summs(ne_hw_82, ne_hw_91)

```

**Plotting Heatwave Characteristic Shifts**

```{r}

####  Plotting shifts using dumbbell plot
hw_dumbbell <- function(wave_summary){
  
  # 1. number of heatwaves
  hw_counts <- ggplot(wave_summary) +
    geom_segment(aes(x = yr, 
                     y = num_waves,
                     xend = yr, 
                     yend = w), 
                 color = "gray60",
                 lineend = "round", 
                 linejoin = "mitre",
                 size = 0.5, 
                 arrow = arrow(length = unit(0.25, "cm"))) +
    geom_point(aes(x = yr, 
                   y = num_waves, 
                   color = "1982-2011 Climatology"), size = 2) +
    geom_point(aes(x = yr,
                   y = w, 
                   color = "1991-2020 Climatology"), size = 2) +
    geom_hline(yintercept = 0, color = "gray30", linetype = 3) +
    labs(y = "# of Events", 
         x = "", 
         color = "") +
    scale_color_gmri() 
  
  # 2. Average duration
  hw_lengths <- ggplot(wave_summary) +
    geom_segment(aes(x = yr, 
                     y = avg_length,
                     xend = yr, 
                     yend = x), 
                 color = "gray60",
                 lineend = "round", 
                 linejoin = "mitre",
                 size = 0.5, 
                 arrow = arrow(length = unit(0.25, "cm"))) +
    geom_point(aes(y = avg_length, 
                   x = yr, 
                   color = "1982-2011 Climatology"), size = 2) +
    geom_point(aes(x = yr, 
                   y = x, 
                   color = "1991-2020 Climatology"), size = 2) +
    geom_hline(yintercept = 0, color = "gray30", linetype = 3) +
    labs(y = "Event Length", 
         x = "", 
         color = "") +
    scale_color_gmri() 
  
  # 3. avg temp
  hw_temps <- ggplot(wave_summary) +
    geom_segment(aes(x = yr, 
                     y = avg_intensity,
                     xend = yr, 
                     yend = y), 
                 color = "gray60",
                 lineend = "round", # See available arrow types in example above
                 linejoin = "mitre",
                 size = 0.5, 
                 arrow = arrow(length = unit(0.25, "cm"))) +
    geom_point(aes(x = yr, 
                   y = avg_intensity, 
                   color = "1982-2011 Climatology"), size = 2) +
    geom_point(aes(x = yr,
                   y = y, 
                   color = "1991-2020 Climatology"), size = 2) +
    geom_hline(yintercept = 0, color = "gray30", linetype = 3) +
    labs(x = "", 
         y = "Avg. Anomaly", 
         color = "") +
    scale_color_gmri() 
  
  # 4. peak temp
  hw_peaks <-  ggplot(wave_summary) +
    geom_segment(aes(x = yr, 
                     y = peak_intensity,
                     xend = yr, 
                     yend = z), 
                 color = "gray60",
                 lineend = "round", 
                 linejoin = "mitre",
                 size = 0.5, 
                 arrow = arrow(length = unit(0.25, "cm"))) +
    geom_point(aes(x = yr,
                   y = peak_intensity, 
                   color = "1982-2011 Climatology"), size = 2) +
    geom_point(aes(x = yr, 
                   y = z,
                   color = "1991-2020 Climatology"), size = 2) +
    geom_hline(yintercept = 0, color = "gray30", linetype = 3) +
    labs(x = "Year", 
         y = "Peak Anomaly",
         color = "") +
    scale_color_gmri() 
  
  # Arrange them
  column_plots <- (hw_counts / hw_lengths / hw_temps / hw_peaks) + plot_layout(guides = "collect") & theme(legend.position = "bottom")
  return(column_plots)

}


```

### Gulf of Maine


```{r, fig.width=8, fig.height=10}
hw_dumbbell(gom_shift) + plot_annotation(title = "Gulf of Maine - Shifting Baselines")
```



### Northeast Shelf

```{r, fig.width=8, fig.height=10}
hw_dumbbell(ne_shift) + plot_annotation(title = "Northeast Shelf - Shifting Baselines")
```




##  Heatwaves that Persist with "New Normal" {.tabset}

What kind of characteristics are shared among the heatwaves that continue to stand out with the new normal.

**Plotting Functions**

```{r}


# Plot the difference within a year
hw_intraannual_plot <- function(hw_data, region){
  
  # Plot how variation happens throughout the year
    hw_data %>%
      ggplot(aes(flat_date, anom_82, color = hwe_change)) +
      geom_point() +
      geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
      scale_color_gmri() +
      labs(x = "", 
           y = "Temperature Anomaly from 82-2011 Climate",
           color = "Impact of New Normal",
           subtitle = str_c(region, " - Change of Heatwave Record with New Normal")) + 
      theme(legend.position = "bottom")

  
}




# Plot seasonal variability
seasonal_polar_plot <- function(hw_data){
  hw_data %>% 
  filter(hwe_change == "Still HW") %>% 
  ggplot(aes(flat_date, anom_82, color = anom_82)) + 
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cc", k = 5),
              color = gmri_cols("orange")) +
  scale_color_distiller(palette = "OrRd", direction = 1) +
  coord_polar() +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0)) +
  theme(panel.border = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom",
        title = element_text(hjust = 0.5)) +
  guides(color = guide_colorbar(title = "Temperature Anomaly from 1982-2011 Climate",
                                title.position = "top", 
                                title.hjust = 0.5,
                                barwidth = unit(4, "inches"), 
                                frame.colour = "black", 
                                ticks.colour = "black")) +
  labs(x = "", y = "", title = "Remaining Heatwaves Under New Climatology")
}


```




###  Gulf of Maine

```{r}
# Summary stats
gom_hw <- gom_comp %>% 
  filter(hwe_82 == TRUE)

# coarse statistics
gom_hw %>% 
  group_by(hwe_change) %>% 
  summarise(n_days = n(),
            avg_anom = round(mean(anom_82),2)) %>% 
  rename("Change in Heatwave Status" = hwe_change, 
         "Number of Days" = n_days, 
         "Average Anomaly" = avg_anom) %>% 
  gt() 

```
  
**Seasonal Patterns in Heatwave Record**  
  
```{r}
# Plot the difference within a year
hw_intraannual_plot(gom_hw, "Gulf of Maine")
```


**Heatwaves that Stand Out in New Normal**

```{r}

# Plot seasonal variability
seasonal_polar_plot(gom_hw)

```

###  Northeast Shelf

```{r}
# Summary stats
shelf_hw <- gom_comp %>% 
  filter(hwe_82 == TRUE)

# coarse statistics
shelf_hw %>% 
  group_by(hwe_change) %>% 
  summarise(n_days = n(),
            avg_anom = round(mean(anom_82),2)) %>% 
  rename("Change in Heatwave Status" = hwe_change, 
         "Number of Days" = n_days, 
         "Average Anomaly" = avg_anom) %>% 
  gt() 


```
  
**Seasonal Patterns in Heatwave Record**  
  
```{r}
# Plot the difference within a year
hw_intraannual_plot(shelf_hw, "Northeast Shelf")
```


**Heatwaves that Stand Out in New Normal**

```{r}

# Plot seasonal variability
seasonal_polar_plot(shelf_hw)

```
