---
title: "OISST Sea Surface Temperature Report"
author: "Adam A. Kemberling"
date: "Updated on: `r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
        collapsed: FALSE
    code_folding: hide
editor_options: 
  chunk_output_type: console
params: 
  region: 
    label: "Select a Focal Region:" 
    value: "Gulf of Maine"
    input: select
    choices: ["Gulf of Maine", "CPR Gulf of Maine", "Northwest Atlantic", "Northeast Shelf"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F, comment = "")

# Packages
library(lubridate)
library(ncdf4)
library(raster)
library(rnaturalearth)
library(sf)
library(stars)
library(gmRi)
library(here)
library(janitor)
library(knitr)
library(patchwork)
library(tidyverse)
library(kableExtra)
library(heatwaveR)
library(plotly)
library(ggpmisc)

# Support Functions
source(here("R/oisst_support_funs.R"))


#box paths
box_paths <- research_access_paths()


# File Paths
mills_path <- box_paths$mills
res_path   <- str_replace(box_paths$res, "RES Data", "RES_Data")
okn_path   <- box_paths$okn  
oisst_path <- paste0(res_path, "OISST/oisst_mainstays/")


# Set ggplot theme for figures
theme_set(theme_bw())


# Set theme up for maps
map_theme <- list(
  theme(
    panel.border       = element_rect(color = "black", fill = NA),
    plot.background    = element_rect(color = "transparent", fill = "transparent"),
    line               = element_blank(),
    axis.title.x       = element_blank(), # turn off titles
    axis.title.y       = element_blank(),
    legend.position    = "bottom", 
    legend.title.align = 0.5))


# Polygons for mapping
new_england <- ne_states("united states of america") %>% st_as_sf(crs = 4326) 
canada      <- ne_states("canada") %>% st_as_sf(crs = 4326)
world       <- ne_countries() %>% st_as_sf(crs = 4326)
greenland   <- ne_states(country = "greenland") %>% st_as_sfc(crs = 4326)
```

`r gmRi::use_gmri_style_rmd(css = "gmri_rmarkdown.css")`


#  `r params$region` Sea Surface Temperature Outlook


## Region Extent {.tabset .tabset-pills}

The spatial extent for `r params$region` is displayed below. This bounding box is the same bounding box coordinates used to clip the OISST data when constructing the time series data from the array.



```{r}
# testing regions
# params$region <- "Northeast Shelf"
# params$region <- "cpr gulf of maine"
# params$region <- "gulf of maine"
# params$region <- "northwest atlantic"

# File paths for various extents based on params$region
poly_path <- switch(
  EXPR = tolower(params$region),
  "gulf of maine"      = "Shapefiles/gmri_sst_focal_areas/apershing_gulf_of_maine.geojson",
  "cpr gulf of maine"  = "Shapefiles/gmri_sst_focal_areas/cpr_gulf_of_maine.geojson",
  "northwest atlantic" = "Shapefiles/gmri_sst_focal_areas/aak_northwest_atlantic.geojson",
  "northeast shelf"    = "Shapefiles/NELME_regions/NELME_sf.shp")


# Load the bounding box for Andy's GOM to show they align
poly_path <- paste0(res_path, poly_path)
region_extent <- st_read(poly_path, quiet = TRUE)


# Pull extents for the region for crop
crop_x <- st_bbox(region_extent)[c(1,3)]
crop_y <- st_bbox(region_extent)[c(2,4)]


# Zoom out for cpr extent
if(tolower(params$region) == "cpr gulf of maine"){
  crop_x <- c(-70.875, -65.375)
  crop_y <- c(40.375,   45.125)}


# one off shapes
# region_extent <- read_sf(paste0(res_path, ""))

# Full plot
ggplot() +
  geom_sf(data = new_england, fill = "gray90") +
  geom_sf(data = canada, fill = "gray90") +
  geom_sf(data = greenland, fill = "gray90") +
  geom_sf(data = region_extent, 
          color = gmri_cols("gmri blue"), 
          fill = gmri_cols("gmri blue"), alpha = 0.2, linetype = 2, size = 0.5) +
  map_theme +
  coord_sf(xlim = crop_x, 
           ylim = crop_y, expand = T) 


```



# GMRI OISST Resources on Box

For many of our most-used areas, regional climatologies and anomalies have been processed using a handful of jupyter notebook work flows to take advantage of {[xarray](http://xarray.pydata.org/en/stable/index.html)}. 

The following sections will highlight what those work flows are, what they detail, and how they can be updated using the one of our focal regions as an example.

The root folder for the following OISST products is `~/Box/RES_Data/OISST/oisst_mainstays`.


##  Regional Timeseries

Area-specific time series are the most basic building block for relaying temporal trends. For any desired area (represented by a spatial polygon) we can generate a time series table of the mean sea surface temperature within that area for each day. Additionally, we can compare how observed temperatures correspond with the expected conditions based on a climatology using a specified  reference period. 

Pre-processed tables have been placed on box for quick access from within the `RES_Data` folder: `RES_Data/OISST/oisst_mainstays/regional_timeseries`.    

A function to quickly access these time-lines has been added to the {[gmRi](https://github.com/gulfofmaine/gmri)} package as `oisst_access_timeseries()`.

Additional functions exists to both look up what timeseries are available `gmRi::get_region_names()`, with additional functionality to return both the path to the data and the path to the shapefile used to generate the timeseries: `gmRi::get_timeseries_paths()`

```{r}

# Use {gmRi} instead to load timeseries to tye up loose ends

#finnicky naming
poly_name <- switch(EXPR = params$region,
                    "Gulf of Maine"      = "apershing gulf of maine", 
                    "CPR Gulf of Maine"  = "cpr gulf of maine", 
                    "Northwest Atlantic" = "northwest atlantic",
                    "Northeast Shelf"    = "northeast us shelf" )




# load with oisst_access_timeseries
if(params$region == "Northeast Shelf"){
  region_timeseries <- read_csv(paste0(oisst_path, "regional_timeseries/NELME_regions/OISSTv2_anom_NELME.csv"))
} else {
  region_timeseries <- oisst_access_timeseries(oisst_path = oisst_path, 
                                               region_family = "gmri focus areas", 
                                               poly_name = poly_name)
}

# format dates
region_timeseries <- region_timeseries %>% 
  mutate(time = as.Date(time))

# Display Table of first 6 entries
head(region_timeseries) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  kable() %>% 
  kable_styling()

# march 1st sst
mar1 <- region_timeseries %>% 
  filter(modified_ordinal_day == 61) %>% 
  distinct(sst_clim) %>% 
  pull(sst_clim)
```


Each of our Climatologies are currently set up to calculate daily averages on a modified year day, such that every March 1st and all days after fall on the same day, regardless of whether it is a leap year or not. 

This preserves comparisons across calendar dates such-as: "The average temperature on march 1st is `r mar1`` for the reference period 1982 to 2011"


In these tables `sst` is the mean temperature observed for that date averaged across all cells within the area. `sst_clim` & `clim_sd` are the climate means and standard deviations for a 1982-2011 climatology. `sst_anom` is the daily observed minus the climate mean.


### Warming Rates {.tabset .tabset-pills}

Regional warming trends below were calculated using all the available data for complete years beginning with 1982 through the end of 2020.


#### Annual

```{r yearly summaries}
# Summarize by year to return mean annual anomalies and variance
annual_summary <- region_timeseries %>% 
  mutate(year = year(time)) %>% 
  group_by(year) %>% 
  summarise(sst = mean(sst, na.rm = T),
            sst_anom = mean(sst_anom, na.rm = T), 
            .groups = "keep") %>% 
  ungroup() %>% 
  mutate(yr_as_dtime = as.Date(paste0(year, "-07-02")))


# # Global Temperature Anomaly Rates
global_anoms <- read_csv(paste0(oisst_path, "global_timeseries/global_anoms_1982to2011.csv"))
global_anoms <- mutate(global_anoms, year = year(time))

global_summary <- global_anoms %>% 
  group_by(year) %>% 
  summarise(sst_anom = mean(sst, na.rm = T), .groups = "keep") %>% 
  ungroup() %>% 
  mutate(yr_as_dtime = as.Date(paste0(year, "-07-02")))
```



```{r annual regressions}

# Build Regression Equation Labels
lm_all <- lm(sst_anom ~ year, 
             data = filter(annual_summary, year %in% c(1982:2020))) %>% 
  coef() %>% 
  round(3)
lm_15  <- lm(sst_anom ~ year, 
             data = filter(annual_summary, year %in% c(2006:2020))) %>% 
  coef() %>% 
  round(3)
lm_global <- lm(sst_anom ~ year, 
                data = filter(global_summary, year %in% c(1982:2020))) %>% 
  coef() %>% 
  round(3)


# Convert yearly rate to decadal
decade_all    <- lm_all['year'] * 10
decade_15     <- lm_15['year'] * 10
decade_global <- lm_global["year"] * 10


# Equation to paste in
eq_all    <- paste0("y = ", lm_all['(Intercept)'], " + ", decade_all, " x")
eq_15     <- paste0("y = ", lm_15['(Intercept)'], " + ", decade_15, " x")
eq_global <- paste0("y = ", lm_15['(Intercept)'], " + ", decade_global, " x")

```





```{r annual trend plot}
####  Annual Trend Plot  ####
ggplot(data = annual_summary, aes(yr_as_dtime, sst_anom)) +
  
  # Add daily data
  geom_line(data = region_timeseries,
            aes(time, sst_anom),
            alpha = 0.5, color = "gray") +
  
  # Overlay yearly means
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.7) +
  
  # Add regression lines 
  geom_smooth(data = filter(global_summary, year <= 2020),
            method = "lm", 
            aes(color = "1982-2020 Global Trend"),
            formula = y ~ x, se = F, 
            linetype = 3) +
  geom_smooth(data = filter(annual_summary, year <= 2020),
              method = "lm",
              aes(color = "1982-2020 Regional Trend"),
              formula = y ~ x, se = F,
              linetype = 2) +
  geom_smooth(data = filter(annual_summary, year %in% c(2006:2020)),
              method = "lm", 
              aes(color = "2006-2020 Regional Trend"),
              formula = y ~ x, se = F, 
              linetype = 2) +

  
  # Manually add equations so they show yearly not daily coeff
  geom_text(data = data.frame(), 
            aes(label = eq_all, x = min(region_timeseries$time), y = Inf),
            hjust = 0, vjust = 2, color = gmri_cols("gmri blue")) +
  geom_text(data = data.frame(), 
            aes(label = eq_15, x = min(region_timeseries$time), y = Inf),
            hjust = 0, vjust = 3.5, color = gmri_cols("orange")) +
  geom_text(data = data.frame(), 
            aes(label = eq_global, x = min(region_timeseries$time), y = Inf),
            hjust = 0, vjust = 5, color = gmri_cols("green")) +

  # Colors
  scale_color_manual(values = c(
    "1982-2020 Regional Trend" = as.character(gmri_cols("gmri blue")),
    "2006-2020 Regional Trend" = as.character(gmri_cols("orange")),
    "1982-2020 Global Trend"   = as.character(gmri_cols("green")))) +
  
  # labels + theme
  labs(x = "", 
       y = expression("Sea Surface Temperature Anomaly"~degree~C),
       caption = paste0("Regression coefficients reflect decadal change in sea surface temperature.
                         Anomalies calculated using 1982-2011 reference period.")) +
   theme(legend.title = element_blank(),
        legend.position = c(0.85, 0.1),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent", color = "transparent"),
        panel.grid = element_blank())


```


#### Quarterly


```{r}
# Repeat for the seasons (equal quarters here*)
quarter_summary <- region_timeseries %>% 
  mutate(year = year(time),
         season = factor(quarter(time, fiscal_start = 1)),
         season = fct_recode(season, 
                             c("Jan 1 - March 31" = "1"), 
                             c("Apr 1 - Jun 30" = "2"), 
                             c("Jul 1 - Sep 30" = "3"), 
                             c("Oct 1 - Dec 31" = "4"))) %>% 
  group_by(year, season) %>% 
  summarise(sst = mean(sst, na.rm = T),
            sst_anom = mean(sst_anom, na.rm = T), 
            .groups = "keep") 

# Plot
quarter_summary %>% 
  ggplot(aes(year, sst_anom)) +
  geom_line(group = 1) +
  geom_point() +
  geom_smooth(method = "lm", 
              aes(color = "Regional Trend"),
              formula = y ~ x, se = F, linetype = 2) +
  stat_poly_eq(formula = y ~ x,
               color = gmri_cols("gmri blue"),
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = T) +
  scale_color_manual(values = c("Regional Trend" = as.character(gmri_cols("orange")))) +
  labs(x = "", 
       y = expression("Sea Surface Temperature Anomaly"~degree~C),
       caption = "Regression coefficients reflect annual change in sea surface temperature.") +
  theme(legend.title = element_blank(),
        legend.position = c(0.875, 0.05),
        legend.background = element_rect(fill = "transparent"),
        legend.key = element_rect(fill = "transparent", color = "transparent")) +
  facet_wrap(~season)
```




### Marine Heatwaves {.tabset .tabset-pills}

For the figures below heatwave events were determined using the methods of Hobday et al. 2016 and implemented using the R package [{heatwaveR}](https://robwschlegel.github.io/heatwaveR/) using a 90% threshold for both heatwave and cold spell events. 


The [{heatwaveR}](https://robwschlegel.github.io/heatwaveR/) package provides a relatively quick way of working with tabular data to calculate a seasonal climate mean, as well as heatwaves and coldwaves at a desired threshold.

These functions can be used to get the climatology using the standard day of year if desired, and track the number and length of heatwave events. Heatwave events follow the definition of [Hobday et al. 2016](http://www.marineheatwaves.org/all-about-mhws.html) and are set up to use a 90% threshold for heatwave/coldwave detection.


> A marine heatwave is defined a when seawater temperatures exceed a seasonally-varying threshold (usually the 90th percentile) for at least 5 consecutive days. Successive heatwaves with gaps of 2 days or less are considered part of the same event.

The heatwave history for the region displayed above is as follows:

```{r}

# Use function to process heatwave data for plotting
region_heatwaves <- pull_heatwave_events(region_timeseries, threshold = 90)

```


Heatwave timelines can be then be plotted using the {ggplot2} package for static plots, or by using the {plotly} package for interactive content.


#### Interactive SST Timeline


For anything we wish to host on the web there is an option to display tables and graphs that are interactive. The {plotly} package is one-such tool for producing plots that allow users to pan, zoom, and highlight discrete observations.



```{r}
# Grab data from the most recent year through present day to plot
last_year <- Sys.Date() - 365
last_year <- last_year - yday(last_year) + 1
last_yr_heatwaves <- region_heatwaves %>% 
  filter(time >= last_year)


# Get number of heatwave events and total heatwave days for last year
# How many heatwave events:
last_full_yr <- last_yr_heatwaves %>% filter(year(time) == year(last_year))
num_events   <- max(last_full_yr$mhw_event_no, na.rm = T) - min(last_full_yr$mhw_event_no, na.rm = T)

# How many heatwave days
num_hw_days <- sum(last_full_yr$mhw_event, na.rm = T)
```



```{r mhw plotly}

# Plot the interactive timeseries
last_yr_heatwaves %>% 
  filter(time >= last_year) %>% 
  plotly_mhw_plots()

```

#### Static SST Timeline



```{r mhw ggplot}

# Set colors by name
color_vals <- c(
  "Sea Surface Temperature" = "royalblue",
  "Heatwave Event"          = "darkred",
  "Cold Spell Event"        = "lightblue",
  "MHW Threshold"           = "coral3",
  #"MHW Threshold"           = "gray30",
  "MCS Threshold"           = "skyblue",
  #"MCS Threshold"           = "gray30",
  "Daily Climatology"    = "gray30")


# Set the label with degree symbol
ylab <- expression("Sea Surface Temperature"~degree~C)



# Plot the last 365 days
ggplot(last_yr_heatwaves, aes(x = time)) +
    geom_segment(aes(x = time, xend = time, y = seas, yend = sst), 
                 color = "royalblue", alpha = 0.25) +
    geom_segment(aes(x = time, xend = time, y = mhw_thresh, yend = hwe), 
                 color = "darkred", alpha = 0.25) +
    geom_line(aes(y = sst, color = "Sea Surface Temperature")) +
    geom_line(aes(y = hwe, color = "Heatwave Event")) +
    geom_line(aes(y = cse, color = "Cold Spell Event")) +
    geom_line(aes(y = mhw_thresh, color = "MHW Threshold"), lty = 3, size = .5) +
    geom_line(aes(y = mcs_thresh, color = "MCS Threshold"), lty = 3, size = .5) +
    geom_line(aes(y = seas, color = "Daily Climatology"), lty = 2, size = .5) +
    scale_color_manual(values = color_vals) +
    scale_x_date(date_labels = "%b", date_breaks = "1 month") +
    theme(legend.title = element_blank(),
          legend.position = "bottom") +
    labs(x = "", 
         y = ylab, 
         caption = paste0("X-axis start date: ", last_year, 
                          "\nNumber of Heatwave events in ", 
                          year(last_year), 
                          ":                  ", 
                          num_events,
                          "\nNumber of Heatwave Days in ", 
                          year(last_year),
                          ":              ", 
                          num_hw_days,
                          "\nClimate reference period :  1982-2011"))
```


#### Heatwave Events

 

```{r}

# Prep the legend title
guide_lab <- expression("Sea Surface Temperature Anomaly"~degree~C)

# Set new axis dimensions, y = year, x = day within year
# use a flate_date so that they don't stair step
base_date <- as.Date("2000-01-01")
grid_data <- region_heatwaves %>% 
  mutate(year = year(time),
         yday = yday(time),
         flat_date = as.Date(yday-1, origin = base_date))


# Set palette limits to center it on 0 with scale_fill_distiller
limit <- max(abs(grid_data$sst_anom)) *c(-1,1)


# Assemble heatmap plot
heatwave_heatmap <- ggplot(grid_data, aes(x = flat_date, y = year)) +
  
  # background box fill for missing dates
  geom_rect(xmin = base_date, xmax = base_date + 365, 
            ymin = min(grid_data$year) - .5, ymax = max(grid_data$year) + .5, 
            fill = "gray75", color = "transparent") +
  
  # tile for sst colors
  geom_tile(aes(fill = sst_anom)) +
  
  # points for heatwave events
  geom_point(data = filter(grid_data, mhw_event == TRUE),
             aes(x = flat_date, y = year), size = .25)  +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", expand = c(0,0)) +
  scale_y_continuous(limits = c(1980.5, 2021.5), expand = c(0,0)) +
  labs(x = "", 
       y = "",
       "\nClimate reference period : 1982-2011") +
  
  #scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
  scale_fill_distiller(palette = "RdBu", na.value = "transparent", 
                       limit = limit) +
  
  #5 inches is default rmarkdown height for barheight
  guides("fill" = guide_colorbar(title = guide_lab, 
                                 title.position = "right", 
                                 title.hjust = 0.5,
                                 barheight = unit(4.8, "inches"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +  
  theme_classic() +
  theme(legend.title = element_text(angle = 90))




#### summary side plots:


# number of heatwaves
# remove NA as a distinct heatwave number
n_waves <- grid_data %>% 
    group_by(year(time), mhw_event_no) %>% 
    summarise(total_days = sum(mhw_event, na.rm = T), .groups = "keep") %>% 
    ungroup() %>% 
    drop_na() %>%
    group_by(`year(time)`) %>% 
    summarise(num_waves = n_distinct(mhw_event_no), .groups = "keep")

# average heatwave duration
wave_len <- grid_data %>% 
  group_by(year(time), mhw_event_no) %>% 
  summarise(total_days = sum(mhw_event, na.rm = T), .groups = "keep") %>% 
  ungroup() %>% 
  drop_na() %>%
  group_by(`year(time)`) %>% 
  summarise(avg_length = mean(total_days), .groups = "keep")


wave_summary <- left_join(n_waves, wave_len)



# Assemble pieces
heatwave_heatmap
```




## Sea Surface Temperature Anomaly Maps {.tabset .tabset-pills}

The 2020 global sea surface temperature anomalies have been loaded and displayed below to visualize how different areas of the ocean experience swings in temperature.

```{r}

# Access information to netcdf on box
nc_year <- "2020"
anom_path <- str_c(oisst_path, "annual_anomalies/1982to2011_climatology/daily_anoms_", nc_year, ".nc")


# Load 2020 as stack
anoms_2020 <- stack(anom_path)


```


### Mean Global

The following code will subset the anomalies for July and plot the average sea surface temperature anomalies for that month:

```{r}
# Get the mean temperature anomalies for July
july_dates <- which(str_sub(names(anoms_2020), 7, 8) == "07")
july_avg   <- mean(anoms_2020[[july_dates]])

# Convert wgs84 raster to stars array
july_st <- st_as_stars(rotate(july_avg))


# Set palette limits to center it on 0 with scale_fill_distiller
limit <- max(abs(july_st[[1]]), na.rm = T) * c(-1,1)

# Plot global map
ggplot() +
  geom_stars(data = july_st) +
  geom_sf(data = world, fill = "gray90") +
  scale_fill_distiller(palette = "RdBu", na.value = "transparent", limit = limit) +
  map_theme +
  coord_sf(expand = FALSE) +
  guides("fill" = guide_colorbar(title = "Average Sea Surface Temperature Anomaly",
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black"))
  
```


### Regional Map

Using the July average sst anomalies we can then clip the data to just the `r params$region` for an aerial view of the region. For this execution the bounding box of lat-lon coordinates from the regional extent will be used.

```{r}

# Clip Raster - Convert to stars
shape_extent <- c(crop_x, crop_y)
region_ras   <- crop(rotate(july_avg), extent(shape_extent))
region_st    <- st_as_stars(region_ras)

# Get crop bounds for coord_sf
crop_x <- st_bbox(region_st)[c(1,3)] 
crop_y <- st_bbox(region_st)[c(2,4)]



# Zoom out some for cpr extent to be the same as Andy's GOM
if(tolower(params$region) == "cpr gulf of maine"){
  crop_x <- c(-71, -65.5)
  crop_y <- c(40.5, 45)}


# Set palette limits to center it on 0 with scale_fill_distiller
limit <- max(abs(region_st[[1]]), na.rm = T) * c(-1,1)

# Plot Region - WGS84 Projection
ggplot() +
  geom_stars(data = region_st) +
  geom_sf(data = new_england, fill = "gray90") +
  geom_sf(data = canada, fill = "gray90") +
  geom_sf(data = greenland, fill = "gray90") +
  scale_fill_distiller(palette = "RdBu", na.value = "transparent", limit = limit) +
  map_theme +
  coord_sf(xlim = crop_x, ylim = crop_y, expand = T) +
  guides("fill" = guide_colorbar(
    title = "Average Sea Surface Temperature Anomaly",
    title.position = "top",
    title.hjust = 0.5,
    barwidth = unit(4, "in"),
    frame.colour = "black",
    ticks.colour = "black"))


```



### Heatwave Progression

Currently in development. Idea is to animate heatwave through the most recent heatwave event.

```{r, eval = T}

# Pull the dates of the most recent heatwave
last_event <- max(region_heatwaves$mhw_event_no, na.rm = T)
last_event_dates <- region_heatwaves %>% 
  filter(mhw_event_no == last_event) %>% 
  pull(time)


# Buffer the dates
event_start <- (min(last_event_dates) - 7)
event_stop  <- max(last_event_dates)
date_seq <- seq.Date(from = event_start,
                     to   = event_stop,
                     by   = 1)


# Load the heatwave dates
data_window <- data.frame(time = c(min(date_seq) , max(date_seq) ),
                          lon  = crop_x,
                          lat  = crop_y)

# Pull data
hw_stack <- oisst_window_load(oisst_path = oisst_path, 
                              data_window = data_window, 
                              anomalies = T)


#drop any empty years that bug in
hw_stack <- hw_stack[map(hw_stack, class) != "character"]
```



```{r, eval = T, animation.hook = 'gifski'}

##### Format the layers and loop through the maps  ####


# Grab only current year, format dates
this_yr   <- stack(hw_stack)
day_count <- length(names(this_yr))
day_labs  <- str_replace_all(names(this_yr), "[.]","-")
day_labs  <- str_replace_all(day_labs, "X", "")
day_count <- c(1:day_count) %>% setNames(day_labs)


# Set palette limits to center it on 0 with scale_fill_distiller
limit <- c(max(values(this_yr), na.rm = T) * -1, 
           max(values(this_yr), na.rm = T) )


# Plot Heatwave 1 day at a time
day_plots <- imap(day_count, function(date_index, date_label) {
  
  # grab dates
  heatwaves_st  <- st_as_stars(this_yr[[date_index]])
  
  # plot every day
  day_plot <- ggplot() +
    geom_stars(data = heatwaves_st) +
    geom_sf(data = new_england, fill = "gray90") +
    geom_sf(data = canada, fill = "gray90") +
    geom_sf(data = greenland, fill = "gray90") +
    scale_fill_distiller(palette = "RdBu", 
                         na.value = "transparent", limit = limit
                         ) +
    map_theme +
    coord_sf(xlim = crop_x, ylim = crop_y, expand = T) +
    guides("fill" = guide_colorbar(
      title = expression("Sea Surface Temperature Anomaly"~degree~C),
      title.position = "top",
      title.hjust = 0.5,
      barwidth = unit(4, "in"),
      frame.colour = "black",
      ticks.colour = "black")) +
    labs(subtitle = paste("Heatwave Period", event_start, "-", event_stop, "\nDisplay Date:", date_label)) 
  
  return(day_plot)
  
  
})


walk(day_plots, print)


```

### Front Progression

Same idea as above but looking at the fronts rather than absolute values

```{r, eval = FALSE}
# remotes::install_github("galuardi/boaR", 
#                         force = T, 
#                         build = T, 
#                         dependencies = "ask",
#                         upgrade = "ask")

library(boaR)

# front test
front_test <- this_yr[[1]]
front_matrix <- as.matrix(front_test)
dim(front_matrix)

# getting coordinates
test_coords <- raster::coordinates(front_test)
xcoords <- sort(unique(test_coords[,"x"]))
ycoords <- sort(unique(test_coords[,"y"]))
rownames(front_matrix) <- ycoords
colnames(front_matrix) <- xcoords
image(front_matrix)


# Getting Belkin O'Reilly Fronts
boaR::boa(lon = test_coords[,"x"], 
          lat = test_coords[,"y"], 
          ingrid = front_matrix, 
          nodata = NA, 
          direction = FALSE)

##########################

# Andrew's code for Fronts


## Land and Biophysical regions
shape_path <- shared.path(os.use = "unix", group = "RES_Data", folder = "Shapefiles")
gom_poly<- st_read(paste0(shape_path, "GulfofMainePhysioRegions/PhysioRegions_WGS84.shp"))
focal_area <- st_bbox(gom_poly)
land <- st_read(paste0(shape_path, "ne_50m_land/ne_50m_land.shp"))

# Collecting variables of interest ----------------------------------------
focal_dates<- as.Date(c("2019-04-01", "2019-09-30"))

## Depth - https://www.ngdc.noaa.gov/mgg/global/global.html
depth_info <- rerddap::info("etopo180")
depth <- rxtracto_3D(dataInfo = depth_info, 
                     parameter = "altitude", 
                     xcoord = focal_area[c(1,3)], 
                     ycoord = focal_area[c(2,4)])

depth_rast<- extracto_to_rast(extracto_obj = depth, variable = "depth")

## SST - Daily MUR SST, Final product (MUR-JPL-L4-GLOB-v4.1): https://podaac.jpl.nasa.gov/dataset/MUR-JPL-L4-GLOB-v4.1?ids=&values=&search=MUR (0.01 deg grid, 1 km)

sst_info <- rerddap::info("nasa_jpl_28d8_bd66_24b1")
sst <- rxtracto_3D(dataInfo = sst_info, 
                   parameter = "analysed_sst", 
                   xcoord = as.numeric(focal_area[c(1,3)]), 
                   ycoord = as.numeric(focal_area[c(2,4)]), 
                   tcoord = focal_dates, tName = "time")


sst_rast <- extracto_to_rast(extracto_obj = sst, 
                             variable = "analysed_sst")
```



## Warming Rates {.tabset .tabset-pills}

Global warming rates are calculated by getting the average temperature for each year across all cells in the data, then using a linear regression for each to determine the annual warming rate in degrees Celsius.

The warming rates are then ranked from high to low to identify areas with the most rapid ocean warming. The following maps display the warming rates and the percentile ranks for the top 20% of warming rates, or rates above the 80th percentile of the data.

```{r}
# Access information to warming rate netcdf files on box
rates_path <- paste0(oisst_path, "warming_rates/annual_warming_rates")

# Percentile cutoff for displaying warming rates on map
percentile_cutoff <- 80

```




### 1982-2011 {.tabset}



```{r}
####  Data Prep  ####


# 1982-2011
rates_stack_all <- stack(str_c(rates_path, "1982to2011.nc"), 
                         varname = "annual_warming_rate")
ranks_stack_all <- stack(str_c(rates_path, "1982to2011.nc"), 
                         varname = "rate_percentile")

# Reclassify to discrete scale
rates_prepped <- reclassify_to_discrete(rates_stack = rates_stack_all, 
                                        ranks_stack = ranks_stack_all,
                                        percentile_cutoff = percentile_cutoff)

# pull out components
rates_st     <- rates_prepped$rates
rates_raw_st <- rates_prepped$rates_raw
ranks_st     <- rates_prepped$ranks

# Get the average warming rate and rank across the area:
rates_crop  <- crop(rotate(rates_stack_all), region_extent)
region_rate <- round(cellStats(rates_crop, mean), 3) 

ranks_crop  <- crop(rotate(ranks_stack_all), region_extent)
region_rank <- round(cellStats(ranks_crop, mean) , 3)

```


The averaged sea surface warming rate percentile for the `r params$region` over the 1982-2011 period was: `r region_rank`. With a regional warming rate of `r region_rate` degrees Celsius per year.
<br/>

#### Global Rates


```{r}

####  Plot Setup  ####

# guide label
guide_lab <- expression("Annual Warming Rate - "~degree~C)

# ggplot using stars
ggplot() +
  geom_stars(data = rates_st) +
  #geom_stars(data = rates_raw_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = guide_lab,
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 1982-2020"))
  
```

#### Global Ranks

```{r}

# ggplot using stars
ggplot() +
  geom_stars(data = ranks_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = "Percentile of Annual Warming Rate",
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 1982-2020"))
  
```

### 2006-2020 {.tabset}

```{r}
####  Data Prep  ####

# 2006-2020
rates_stack_all <- stack(str_c(rates_path, "2006to2020.nc"), 
                         varname = "annual_warming_rate")
ranks_stack_all <- stack(str_c(rates_path, "2006to2020.nc"), 
                         varname = "rate_percentile")

rates_prepped <- reclassify_to_discrete(rates_stack = rates_stack_all, 
                                        ranks_stack = ranks_stack_all)

rates_st <- rates_prepped$rates
ranks_st <- rates_prepped$ranks


# Get the average warming rate and rank across the area:
rates_crop <- crop(rotate(rates_stack_all), region_extent)
region_rate <- round(cellStats(rates_crop, mean), 3) 

ranks_crop <- crop(rotate(ranks_stack_all), region_extent)
region_rank <- round(cellStats(ranks_crop, mean) , 3)

```


The averaged sea surface warming rate percentile for the `r params$region` over the 2006-2020 period was: `r region_rank`. With a regional warming rate of `r region_rate` degrees Celsius per year.   
<br/>

#### Global Rates


```{r}
####  Plot Setup  ####

# guide label
guide_lab <- expression("Annual Warming Rate - "~degree~C)

# ggplot using stars
ggplot() +
  geom_stars(data = rates_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = guide_lab,
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 2006-2020"))
  
```

#### Global Ranks

```{r}
# ggplot using stars
ggplot() +
  geom_stars(data = ranks_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = "Percentile of Annual Warming Rate",
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 2006-2020"))
  
```

### 1982-2020 {.tabset}

```{r}
####  Data Prep  ####


# 1982-2020
rates_stack_all <- stack(str_c(rates_path, "1982to2020.nc"), 
                         varname = "annual_warming_rate")
ranks_stack_all <- stack(str_c(rates_path, "1982to2020.nc"), 
                         varname = "rate_percentile")

rates_prepped <- reclassify_to_discrete(rates_stack = rates_stack_all, 
                                        ranks_stack = ranks_stack_all,
                                        percentile_cutoff = percentile_cutoff)

rates_st <- rates_prepped$rates
rates_raw_st <- rates_prepped$rates_raw
ranks_st <- rates_prepped$ranks


# Get the average warming rate and rank across the area:
rates_crop <- crop(rotate(rates_stack_all), region_extent)
region_rate <- round(cellStats(rates_crop, mean), 3) 

ranks_crop <- crop(rotate(ranks_stack_all), region_extent)
region_rank <- round(cellStats(ranks_crop, mean) , 3)

```


The averaged sea surface warming rate percentile for the `r params$region` over the 1982-2020 period was: `r region_rank`. With a regional warming rate of `r region_rate` degrees Celsius per year.
<br/>

#### Global Rates


```{r}

####  Plot Setup  ####

# guide label
guide_lab <- expression("Annual Warming Rate - "~degree~C)

# ggplot using stars
ggplot() +
  geom_stars(data = rates_st) +
  #geom_stars(data = rates_raw_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = guide_lab,
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 1982-2020"))
  
```

#### Global Ranks

```{r}

# ggplot using stars
ggplot() +
  geom_stars(data = ranks_st) +
  geom_sf(data = world, fill = "gray90") + 
  scale_fill_viridis_c(na.value = "black", option = "plasma") +
  guides("fill" = guide_colorbar(title = "Percentile of Annual Warming Rate",
                                 title.position = "top", 
                                 title.hjust = 0.5,
                                 barwidth = unit(4, "in"), 
                                 frame.colour = "black", 
                                 ticks.colour = "black")) +
  map_theme +
  coord_sf(expand = FALSE) +
  labs(caption = str_c("Top ", percentile_cutoff, "% of Warming Rates Shown.",
                       "\n Data from Years: 1982-2020"))
  
```

`r insert_gmri_footer()`
